---
title: Homework 1
author: Ruby He 
date: "2025-09-12"
output: github_document
---

Load libraries
```{r}
library(moderndive)
library(tidyverse)
```

## Problem 1

```{r}
data(early_january_weather) 
```

The variables in this dataset are `r names(early_january_weather)`.

This dataset has `r ncol(early_january_weather)` variables and `r nrow(early_january_weather)` observations. 

The mean temperature is `r mean(early_january_weather$temp)`.

Scatterplot
```{r}
ggplot(early_january_weather, aes(x = time_hour, y = temp, color = humid)) + geom_point()
ggsave("scatter_plot.pdf",height = 4, width = 6)
```
In the scatterplot we can see that temperature rises as the time increases. We can also see that humidity also seems to increase as the temperature rises as well. 

## Problem 2

```{r}
prob2_df=
  tibble(
    num_samp = rnorm (10),
    greater_0 = num_samp > 0,
    char_vec = c("x","y","z","a","b","c","d","e","f","g"),
    fac_vec = factor(c("low","medium","low","high","high","low","low","medium","high","low"))
  )

mean(pull(prob2_df,num_samp))
mean(pull(prob2_df,greater_0))
mean(pull(prob2_df,char_vec))
mean(pull(prob2_df,fac_vec))
```

you are unable to take the mean of the character and factor vectors, only the numeric and logical vectors produced a mean. 

```{r results = "hide"}
as.numeric(prob2_df$greater_0)
as.numeric(prob2_df$char_vec)
as.numeric(prob2_df$fac_vec)
```

When you convert the logical, character, and factor variables to numeric variables, the logical vector becomes all 1s and 0s. The character vector becomes all NA. And the factor vector reassigns the levels to 1,2,3. 

This would explain why when we try to take the mean prior to using as.numeric, the logical vector is able to produce a mean but the character and vector are not able to. Even if we do convert it to numeric, the characters are unable to become numbers and the factor as a number will not be helpful in finding a mean. 
